Empirical Bayes estimation of the normal inverse Gaussian prior in simultaneous-equation models

### 2019-06-04 docs/manuscript1.Rnw
commit before push
first ENIG simulations

	\section{Model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	We have continuous measurements $y_{id}$ for observation $i=1,\dots, n$ on 
	outcome $d=1,\dots,D$. Throughout this paper we assume the $y_{id}$ to be 
	centred per outcome and let $\y_d = \begin{bmatrix} y_{1d} & \dots & y_{nd} 
	\end{bmatrix} \tr$. We predict the $y_{id}$ with features $x_{ij}$, 
	$j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} 
	\end{bmatrix} \tr$. We have characteristics of the outcomes available in the 
	form of `co-data' $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & 
	\mathbf{c}_D \end{bmatrix} \tr$.

@@

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with the following priors:
	\begin{subequations}\label{eq:priormodel}
	  \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
      \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim 
      \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \label{eq:betaprior} \\
      \gamma_d^2 & \sim \mathcal{IG}(\theta_d, \lambda_d), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG}(\theta, \lambda)$ denotes the inverse Gaussian 
  distribution with mean $\theta$ and shape $\lambda$.
  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_d^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_d$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). See SM Section 
  \ref{sm-sec:nigprior} for more details on this prior.
  In Figure \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution marginalised over $\gamma_d^2$ is depicted together with 
  three other common prior choices,
  Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), ridge and lasso, all scaled to variance one. From this Figure 
  we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_d$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean $\theta_d$ as a 
  function of the co-data $\mathbf{c}_d$ more conveniently than the Student's 
  $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="80%", fig.asp=2/3>>=
library(GeneralizedHyperbolic)
library(metRology)
# all scaled such that variance is one
sigma <- 1
theta <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
eta <- lambda/theta + 2
col <- c(1:6)
lty <- c(1:6)
labels <- as.expression(c(bquote("NIG, "~lambda==.(lambda[1])~", "~
                                   theta==.(theta)),
                          bquote("NIG, "~lambda==.(lambda[2])~", "~
                                   theta==.(theta)),
                          bquote("Student's t, "~lambda==.(lambda[1])~", "~
                                   eta==.(eta[1])),
                          bquote("Student's t, "~lambda==.(lambda[2])~", "~
                                   eta==.(eta[2])),
                          "ridge", "lasso"))
dprior <- function(x, lambda, theta, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(theta*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

ylim <- c(0, dt.scaled(0, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])))
curve(dprior(x, lambda[1], theta, sigma), -3, 3, ylab=expression(p(beta[j])), 
      xlab=expression(beta[j]), n=1000, ylim=ylim,
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], theta, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
curve(dt.scaled(x, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])), add=TRUE, 
      n=1000, col=col[5], lty=lty[5])
curve(dt.scaled(x, eta[2]/2, 0, sqrt(lambda[2]*sigma^2/eta[2])), add=TRUE, 
      n=1000, col=col[6], lty=lty[6])
legend("topright", legend=labels, col=col, lty=lty, 
       title="Prior", seg.len=1)
@

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance.

  - Jeffrey's prior for variance $1/\sigma^{3/2}$ (a priori independent data 
	mean and variance) or $1/\sigma^3$ (joint prior, but only mean and variance, 
	not variance and betas).

	\section{Estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_d^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot q(\gamma_d^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma^2_d) & \overset{D}{=} \mathcal{GIG}
    \(-\frac{p+1}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
    \delta_d^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_d) \tr \bmu^{(h+1)}_d +
    \trace (\bSigma_d^{(h+1)})\] + \lambda_d, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + 
    b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}
    {\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p - 1}{2}} 
    \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}
    {K_{\frac{p+1}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + 
    \frac{p+1}{\delta_d^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of such 
  functions as in (\ref{eq:ratiomodifiedbessel}) is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

  \subsection{Empirical Bayes}\label{sec:empiricalbayes}
  A full Bayesian model requires the specification of hyper parameters 
  $\bm{\theta}=\begin{bmatrix} \theta_1 & \cdots & \theta_D \end{bmatrix} \tr$ 
  and $\begin{bmatrix} \lambda_1 & \cdots & \lambda_D \end{bmatrix} \tr$. 
  These are abstract and hard to interpret parameters 
  for which we generally lack expert knowledge. They do, however, have a 
  significant influence on the shape of the posterior distribution. We 
  therefore propose to estimate these hyper parameters by empirical Bayes. 
  Simply put, empirical Bayes fits the prior model to the data and is, as such, 
  objective (up to model specification).

	The canonical method for empirical Bayes is maximisation of the marginal 
	likelihood with respect to the hyper parameters. In 
	\cite{casella_empirical_2001} the marginal likelihood is maximised by an EM 
	algorithm:
	$$
	\bm{\theta}^{(l+1)},\bm{\lambda}^{(l+1)} = \underset{\bm{\theta},\bm{\lambda}}
	{\argmax}\E_{\cdot | \Y} 
	[\log p(\Y, \B, \bgamma^2, \bsigma^2) | \bm{\theta}^{(l)},\bm{\lambda}^{(l)}],
	$$
	where $\Y = \begin{bmatrix} \y_1 & \cdots & \y_D \end{bmatrix}$,
	$\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$, 
	$\bgamma^2 = \begin{bmatrix} \gamma^2_1 & \cdots & \gamma^2_D 
	\end{bmatrix}\tr$, $\bsigma^2 = \begin{bmatrix} \sigma^2_1 & \cdots & 
	\sigma^2_D \end{bmatrix} \tr$,
	and the expectation is with respect to the posterior. In our case, this 
	posterior is not available in closed form, which renders the expectation 
	difficult. While \cite{casella_empirical_2001} suggests to approximate the 
	expectation by a Monte Carlo sample, we propose to use the variational Bayes 
	approximation developed in Section \ref{sec:variationalbayes}:
	\begin{align*}
	  \bm{\theta}^{(l+1)},\bm{\lambda}^{(l+1)} & = 
	  \underset{\bm{\theta},\bm{\lambda}}{\argmax}\E_{Q^{(l)}} 
	  [\log p(\Y, \B, \bgamma^2, \bsigma^2)] \\
	  & = \underset{\bm{\theta},\bm{\lambda}}{\argmax} \E_{Q^{(l)}} 
	  [\log \pi (\B | \bgamma^2, \bsigma^2)] + 
	  \E_{Q^{(l)}} [\log \pi (\bgamma^2)].
	\end{align*}
	where now the expectation is with respect to the converged variational 
	posterior $Q^{(l)}=\prod_{d=1}^D Q_d^{(l)}$.

	To make use of the outcome-specific codata, we parametrise 
	$\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ and estimate $\blambda$ and
	$\bm{\alpha}$. This allows us to include continuous co-data on the outcomes
	in an objective, data-driven manner. Additionally, it reduces the number of
	hyperparameters even further. The estimating equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \diag 
    (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag 
    (\lambda_d) \mathbf{1}_{D \times 1}, \\
    \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 
    2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - 
  (p + 1)/\delta_d^{(l)})] \cdot \delta_d^{(l)} 
  (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by 
  iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on 
  predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$.

@@

  \section{General model}
  In addition to the outcome co-data $\mathbf{c}_d$, we may have feature co-data
  $\mathbf{z}_j$ available. Examples of this in omics settings are pathway 
  memberships or previously obtained $p$-values for the features. In
  that case, we extend model (\ref{eq:priormodel}) to:
  \begin{subequations}\label{eq:generalmodel}
    \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \nonumber \\
      \beta_{jd} | \tau_j^2, \gamma_d^2, \sigma_d^2 & \sim \mathcal{N} 
      (0, \sigma_d^2 \gamma_d^2 \tau_j^2), \\
      \gamma_d^2 & \sim \mathcal{IG}\(\theta_d^c, \lambda^c_d\), \\
      \tau_j^2 & \sim \mathcal{IG}\(\theta_j^z, \lambda^z_j\), \\
      \sigma_d^2 & \sim 1/\sigma_d^3, \nonumber
    \end{align}
  \end{subequations}
  where we now have an extra variance component $\tau_j$ that is 
  \textit{a priori} inverse Gaussian distributed with mean dependent on the 
  feature co-data $\mathbf{z}_j$. We will refer to this model as the normal
  inverse Gaussian-inverse Gaussian (NIGIG) model. See SM Section 
  \ref{sm-sec:nigigprior} for more details on this prior.

>>>>

	\section{Model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	- change the name to seemingly unrelated regressions?

	We have continuous measurements $y_{id}$ for observation $i=1,\dots, n$ on 
	outcome $d=1,\dots,D$. Throughout this paper we assume the $y_{id}$ to be 
	centred per outcome and let $\y_d = \begin{bmatrix} y_{1d} & \dots & y_{nd} 
	\end{bmatrix} \tr$. We predict the $y_{id}$ with features $x_{ij}$, 
	$j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} 
	\end{bmatrix} \tr$. We have characteristics of the outcomes available in the 
	form of `co-data' $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & 
	\mathbf{c}_D \end{bmatrix} \tr$.

@@

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with the following priors:
	\begin{subequations}\label{eq:priormodel}
	  \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
      \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim 
      \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \label{eq:betaprior} \\
      \gamma_d^2 & \sim \mathcal{IG} \( (\mathbf{c}_d \tr \bm{\alpha})^{-1}, 
      \lambda_d \), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG}((\mathbf{c}_d \tr \bm{\alpha})^{-1}, \lambda_d)$ 
  denotes the inverse Gaussian distribution with mean 
  $(\mathbf{c}_d \tr \bm{\alpha})^{-1}$ and shape $\lambda_d$.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_d$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). See SM Section 
  \ref{sm-sec:nigprior} for more details on this prior.
  In Figure \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution marginalised over $\gamma_d^2$ is depicted together with 
  three other common prior choices,
  Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), normal (ridge) and Laplace (lasso), all scaled to variance one. 
  From this Figure we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_d$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean as a 
  function of the co-data $\mathbf{c}_d$ more conveniently than the Student's 
  $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="80%", fig.asp=2/3>>=
library(GeneralizedHyperbolic)
library(metRology)
# all scaled such that variance is one
sigma <- 1
mgammasq <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
eta <- lambda/mgammasq + 2
col <- c(1:6)
lty <- c(1:6)
labels <- as.expression(c(bquote("NIG, "~lambda==.(lambda[1])~", "~
                                   (bold(c)^"T"~bold(alpha))^-1==.(mgammasq)),
                          bquote("NIG, "~lambda==.(lambda[2])~", "~
                                   (bold(c)^"T"~bold(alpha))^-1==.(mgammasq)),
                          bquote("Student's t, "~lambda==.(lambda[1])~", "~
                                   eta==.(eta[1])),
                          bquote("Student's t, "~lambda==.(lambda[2])~", "~
                                   eta==.(eta[2])),
                          "ridge", "lasso"))
dprior <- function(x, lambda, mgammasq, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(mgammasq*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

ylim <- c(0, dt.scaled(0, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])))
opar <- par(no.readonly=TRUE)
par(mar=opar$mar*c(1, 1.1, 1, 1))
curve(dprior(x, lambda[1], mgammasq, sigma), -3, 3, 
      ylab=expression(p(beta[j])), 
      xlab=expression(beta[j]), n=1000, ylim=ylim,
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], mgammasq, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
curve(dt.scaled(x, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])), add=TRUE, 
      n=1000, col=col[5], lty=lty[5])
curve(dt.scaled(x, eta[2]/2, 0, sqrt(lambda[2]*sigma^2/eta[2])), add=TRUE, 
      n=1000, col=col[6], lty=lty[6])
legend("topright", legend=labels, col=col, lty=lty, 
       title="Prior", seg.len=1)
par(opar)
@

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_d^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response.

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance. This
  joint consideration results in the somewhat unorthodox $1/\sigma^3$ Jeffrey's
  prior.

	\section{Estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_d^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot q(\gamma_d^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma^2_d) & \overset{D}{=} \mathcal{GIG}
    \(-\frac{p+1}{2}, \lambda_d (\mathbf{c}_d \tr \bm{\alpha})^2, \delta_d\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
    \delta_d^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_d) \tr \bmu^{(h+1)}_d +
    \trace (\bSigma_d^{(h+1)})\] + \lambda_d, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + 
    b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = 
    \sqrt{\frac{\lambda_d(\mathbf{c}_d \tr \bm{\alpha})^{2}}
    {\delta_d^{(h)}}} \frac{K_{\frac{p - 1}{2}} 
    \( \sqrt{\lambda_d \delta_d^{(h)}}(\mathbf{c}_d \tr \bm{\alpha}) \)}
    {K_{\frac{p+1}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}
    (\mathbf{c}_d \tr \bm{\alpha}) \)} + 
    \frac{p+1}{\delta_d^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of such 
  functions as in (\ref{eq:ratiomodifiedbessel}) is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

  \subsection{Empirical Bayes}\label{sec:empiricalbayes}
  To make use of the outcome-specific codata, we parametrised the prior 
  $\gamma_d^2$ mean as $(\mathbf{c}_d \tr \bm{\alpha})^{-1}$. This allows us to
  include continuous co-data on the outcomes into the model. Additionally, it 
  reduces the number of hyperparameters. A full Bayesian model then requires the
  specification of these hyper parameters 
  $\bm{\alpha}$ and $\begin{bmatrix} \lambda_1 & \cdots & \lambda_D 
  \end{bmatrix} \tr$. These are abstract and hard to interpret parameters 
  for which we generally lack expert knowledge. They do, however, have a 
  significant influence on the shape of the posterior distribution. We 
  therefore propose to estimate these hyper parameters by empirical Bayes. 
  Simply put, empirical Bayes fits the prior model to the data and is, as such, 
  objective (up to model specification). In our case, this results in an   
  objective and data-driven inclusion of the co-data.

	The canonical method for empirical Bayes is maximisation of the marginal 
	likelihood with respect to the hyper parameters. In 
	\cite{casella_empirical_2001} the marginal likelihood is maximised by an EM 
	algorithm:
	$$
	\bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} = \underset{\bm{\alpha},\bm{\lambda}}
	{\argmax}\E_{\cdot | \Y} 
	[\log p(\Y, \B, \bgamma^2, \bsigma^2) | \bm{\alpha}^{(l)},\bm{\lambda}^{(l)}],
	$$
	where $\Y = \begin{bmatrix} \y_1 & \cdots & \y_D \end{bmatrix}$,
	$\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$, 
	$\bgamma^2 = \begin{bmatrix} \gamma^2_1 & \cdots & \gamma^2_D 
	\end{bmatrix}\tr$, $\bsigma^2 = \begin{bmatrix} \sigma^2_1 & \cdots & 
	\sigma^2_D \end{bmatrix} \tr$,
	and the expectation is with respect to the posterior. In our case, this 
	posterior is not available in closed form, which renders the expectation 
	difficult. While \cite{casella_empirical_2001} suggests to approximate the 
	expectation by a Monte Carlo sample, we propose to use the variational Bayes 
	approximation developed in Section \ref{sec:variationalbayes}:
	\begin{align*}
	  \bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} & = 
	  \underset{\bm{\alpha},\bm{\lambda}}{\argmax}\E_{Q^{(l)}} 
	  [\log p(\Y, \B, \bgamma^2, \bsigma^2)] \\
	  & = \underset{\bm{\alpha},\bm{\lambda}}{\argmax} 
	  \E_{Q^{(l)}} [\log \pi (\bgamma^2)].
	\end{align*}
	where now the expectation is with respect to the converged variational 
	posterior $Q^{(l)}=\prod_{d=1}^D Q_d^{(l)}$.

	The estimating equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \diag 
    (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag 
    (\lambda_d) \mathbf{1}_{D \times 1}, \\
    \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 
    2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - 
  (p + 1)/\delta_d^{(l)})] \cdot \delta_d^{(l)}
  /[\lambda_d^{(l)}(\mathbf{c}_d \tr \alpha^{(h)})^2]$. 
  Solving the equations is done by 
  iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on 
  predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$.

@@

  \section{Extended model I}
  In addition to the outcome co-data $\mathbf{c}_d$, we may have feature co-data
  $\mathbf{z}_j$ available. Examples of this in omics settings are pathway 
  memberships or previously obtained $p$-values for the features. In
  that case, we extend model (\ref{eq:priormodel}) to:
  \begin{subequations}\label{eq:extendedmodel1}
    \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \nonumber \\
      \beta_{jd} | \tau_j^2, \gamma_d^2, \sigma_d^2 & \sim \mathcal{N} 
      (0, \sigma_d^2 \gamma_d^2 \tau_j^2), \\
      \gamma_d^2 & \sim \mathcal{IG}
      \((\mathbf{c}_d \tr \bm{\alpha})^{-1}, \lambda_d\), \\
      \tau_j^2 & \sim \mathcal{IG}
      \((\mathbf{z}_j \tr \bm{\theta})^{-1}, \kappa_j\), \\
      \sigma_d^2 & \sim 1/\sigma_d^3, \nonumber
    \end{align}
  \end{subequations}
  where we now have an extra variance component $\tau_j$ that is 
  \textit{a priori} inverse Gaussian distributed with mean dependent on the 
  feature co-data $\mathbf{z}_j$ and hyper parameter $\bm{\theta}$. We will
  refer to this model as the double normal
  inverse Gaussian (DNIG) model. See SM Section 
  \ref{sm-sec:dnigprior} for more details on this prior.
